Scattering from compact objects: \\ Regge poles and the Complex Angular Momentum method

### 2019-12-22 rainbowCAM.tex
Small changes in appendix and Sec III.C

The CAM theory was originally developed to deal theoretically with the propagation and diffraction of radio waves around the Earth, by Watson~\cite{Watson18} (see also the work of Sommerfeld~\cite{Sommerfeld49}). It has since been extensively used in several domains of physics involving resonant scattering theory (see, e.g., Refs.~\cite{deAlfaro:1965zz,Newton:1982qc,Watson18,Sommerfeld49,Nussenzveig:2006,Grandy2000,Uberall1992,AkiRichards2002,AkiRichards2002,Gribov69,
Collins77,BaronePredazzi2002,DonnachieETAL2005} as well as references therein for various applications in quantum mechanics, nuclear physics, electromagnetism, optics, seismology and high energy physics). Since the pioneering work in 1994 by Andersson and Thylwe~\cite{Andersson:1994rk,Andersson:1994rm} the CAM theory has been successfully applied to black hole scattering scenarios. They described the scattering of scalar waves by a Schwarzschild black hole using CAM techniques, and showed that, in this case, the Regge poles are associated with ``surface waves'' localised near the unstable light ring at $ r = 3M $ (\textit{i.e.}, the so-called photon sphere). Later, Decanini, Folacci and Jensen showed that the complex frequencies of weakly damped QNMs are Breit-Wigner resonances generated by the surface waves previously mentioned and, using the concept of Regge trajectories, they were able to construct semiclassically the spectrum of the QNM complex frequencies~\cite{Decanini:2002ha}. Recently, Folacci and Ould~El~Hadj have shown that CAM machinery can be used for precise numerical calculations of scattering cross sections of scalar fields, electromagnetic fields~\cite{Folacci:2019cmc} and gravitational waves~\cite{Folacci:2019vtt} on a Schwarzschild space-time. Moreover, using the third WKB approximation to obtain the asymptotic expressions for the lowest Regge poles~\cite{Decanini:2009mu} and associated residues, they have been able to provide an analytical approximation describing accurately both the ``glory'' and a large part of the ``orbiting'' oscillations in black hole scattering cross sections in the short-wavelength regime. An accurate approximation for the black hole absorption cross section has also been derived via an analysis of Regge poles~\cite{Decanini:2011xi,Decanini:2011xw}. In addition, an alternative description of gravitational radiation from black holes based on CAM theory was developed in Ref.~\cite{Folacci:2018sef}.

@@

There is a close relationship between Regge poles and quasi-normal mode (QNM) frequencies; both are sets of poles of the scattering matrix, with the latter lying in the complex-frequency plane. The excitation of black hole QNMs leads to distinctive signatures in \emph{time-dependent} scattering scenarios. For example, in the immediate aftermath of a black hole merger, the perturbed black hole returns to a quiescent state via a ``ringdown phase'' in which its gravitational-wave signal is well described as a sum of quasinormal modes \cite{Giesler:2019uxc}. The QNM spectrum of compact bodies have been well-studied by several authors  \cite{Detweiler:1985zz, Kokkotas:1986gd, Chandrasekhar449, Kokkotas:1992ka, Leins:1993zz, Andersson:1995ez, Andersson1996}, and are reviewed in Ref.~\cite{Kokkotas:1999bd}.

@@

%Finally we apply the CAM-approach developed for black-hole scattering in \cite{Folacci:2019cmc} and \cite{Folacci:2019vtt}.

@@

Furthermore, since the wave equation is invariant under the transformation $\lambda \rightarrow -\lambda$, it follows that the distribution of poles in the left-half of the complex plane follows by reflection through the origin. Thus, for $\omega > 0$, the 4th and 2nd quadrants are devoid of poles, and the 3rd quadrant has the same structure as shown in Fig.~\ref{fig:rp1} and~\ref{fig:rp2}.
Conversely, for $\omega < 0$, we may use the symmetry $\lambda_n(-\omega) = \lambda_n(\omega)^*$ to establish that there are poles in the 2nd and 4th quadrants only.

@@

The Regge pole spectrum for a compact body is qualitatively similar to the Regge pole spectrum found in Mie scattering of electromagnetic waves by a transparent sphere of refractive index $\tilde{n}$. This has been studied since the 1960s; see for example Fig.~9.2 in Ref.~\cite{Nussenzveig:2006}. % In the latter case, the poles are given by the zeros of a combination of special functions.
Henceforth we shall adopt the terminology of Nussenzveig \cite{Nussenzveig:2006}, in which the three branches are labelled as:
\begin{enumerate}
 \item \emph{Broad resonances}: approximately uniformly-spaced poles above the real axis with approximately constant imaginary part; somewhat sensitive to internal structure ($r < R$);
 \item \emph{Surface waves}: highly-damped modes that are relatively insensitive to the internal structure and which depend chiefly on the surface geometry.
 \item \emph{Narrow resonances}: modes approaching the real axis corresponding to trapped modes which only appear in the UCO case ($R < 3M$).
\end{enumerate}
In Figs.~\ref{fig:rp1} and \ref{fig:rp2} the broad resonances, surface waves and narrow resonances are indicated by red squares, blue circles and purple triangles, respectively; and black diamonds indicate the black hole RPs. Whereas there exists an infinite number of poles of the surface-wave branch, in principle, there are no narrow-resonance poles at all in the $R/M = 6$ case, and just $\le 4$ narrow-resonance poles in the $R/M = 2.26$ case. The narrow-resonance branch is observed to end close to the start of the black-hole branch.

@@

%\subsubsection{Regge poles and associated residues}
%\label{SecIIc_2}

%It is important to note that the poles of  $S_{\lambda -1/2} (\omega)$ in the complex $\lambda$ plan (i.e., the Regge poles) are defined  as  the zeros of the coefficient  $A^{(-)}_{\lambda-1/2} (\omega)$ [see Eq.~(\ref{Matrix_S_CAM})], i.e., the values $\lambda_n(\omega)$ such that
%\begin{equation}\label{PR_def_Am}
%A^{(-)}_{\lambda_n(\omega)-1/2} (\omega)=0,
%\end{equation}
%with $n=1,2,3,\ldots$
%The residue of the matrix $S_{\lambda-1/2}(\omega)$ at the pole $\lambda=\lambda_n(\omega)$ is defined by [see Eq.~(\ref{Matrix_S_CAM})]
%\begin{equation}\label{residues_RP}
%r_n(\omega)=e^{i\pi [\lambda_n(\omega)+1/2]} \left[ \frac{A_{\lambda -1/2}^{(+)}(\omega)}{\frac{d}{d \lambda}A_{\lambda -1/2}^{(-)}(\omega)}\right]_{\lambda=\lambda_n(\omega)}.
%\end{equation}
%These residues play a central role in the complex angular momentum paradigm.
%

@@

To construct the scattering amplitude \eqref{Scalar_Scattering_amp}, the background integrals~\eqref{CAM_Scalar_Scattering_amp_decomp_Background_a} and~\eqref{CAM_Scalar_Scattering_amp_decomp_Background_b}, and the Regge pole amplitude~\eqref{CAM_Scalar_Scattering_amp_decomp_RP}, we use, \textit{mutatis mutandis}, the computational methods of Refs~\cite{Folacci:2019cmc,Folacci:2019vtt}. In these works the authors calculated the CAM representation of scattering amplitudes for scalar, electromagnetic and gravitational waves by Schwarzschild BH (see also Ref~\cite{Dolan:2017rtj}).

Due to the long rang nature of the field propagating on the Schwarzschild BH (outside the compact body), the scattering amplitude~\eqref{Scalar_Scattering_amp} and the background integral~\eqref{CAM_Scalar_Scattering_amp_decomp_Background_a} both suffer a lack of convergence. To overcome this problem, \textit{i.e.}, to accelerate the convergence of this sum and integral, we have used the method described in the Appendix of Ref~\cite{Folacci:2019cmc}. We have performed all the numerical calculations by using {\it Mathematica}. % \cite{Mathematica}.

@@

The CAM toolkit was applied here to gain a complementary method for calculating scattering amplitudes. In summary, we have shown that, for the intermediate and high angles, these scattering amplitudes can be reconstructed in terms of Regge poles with great precision in the intermediate and high frequency regimes. For compact objects ($R/M\sim6$), we observed that it is necessary to take into account the background integral contributions to describe the glory, and this is true regardless of the frequency. Conversely, this is not the case for UCOs ($R/M < 3$), where the sum over the Regge poles is sufficient to accurately describe the glory. Moreover, the sum over the Regge poles permitted us to overcome the difficulties linked to the lack of convergence from which the scattering amplitude suffers (\textit{i.e.} partial wave expansions) due to the long-range of the field propagating on the Schwarzschild spacetime outside the compact body. However, the existence of the broad resonance branch %(see also narrow resonance branch for the UCOs)
close to the real axis with an approximately uniform spacing, $\Delta\lambda\sim 2$, requires the calculation of a large number of poles. In the black-hole case, by contrast, there is only a single branch of surface-wave-type modes (see Fig.~\ref{RP_R6_approx_2Mw_3_6_s_1}), and in the high-frequency limit, a few poles ($<5$) capture both the orbiting and glory phenomena \cite{Folacci:2019cmc, Folacci:2019vtt}.

@@

As discussed in Sec.~\ref{subsec:results1}, a good analytic extension of the S-matrix into the left-half complex $\lambda$-plane leads to the symmetry \eqref{Matrix_S_CAM_symm}. Then, it follows that the Regge poles with $\text{Re}[\lambda] < 0 $ may be found for
%
\begin{equation}
    \begin{dcases}
        \omega>0 \\
        \text{Im}[\lambda]<0
     \end{dcases}
\quad\text{or}\quad
    \begin{dcases}
      \omega<0 \\
        \text{Im}[\lambda]>0
    \end{dcases}.
\end{equation}\

>>>>

The CAM theory was originally developed to deal theoretically with the propagation and diffraction of radio waves around the Earth, by Watson~\cite{Watson18} (see also the work of Sommerfeld~\cite{Sommerfeld49}). It has since been extensively used in several domains of physics involving resonant scattering theory (see, e.g., Refs.~\cite{deAlfaro:1965zz,Newton:1982qc,Watson18,Sommerfeld49,Nussenzveig:2006,Grandy2000,Uberall1992,AkiRichards2002,AkiRichards2002,Gribov69,
Collins77,BaronePredazzi2002,DonnachieETAL2005} as well as references therein for various applications in quantum mechanics, nuclear physics, electromagnetism, optics, seismology and high energy physics). Since the pioneering work in 1994 by Andersson and Thylwe~\cite{Andersson:1994rk,Andersson:1994rm} the CAM theory has been successfully applied to black hole scattering scenarios. They described the scattering of scalar waves by a Schwarzschild black hole using CAM techniques, and showed that, in this case, the Regge poles are associated with ``surface waves'' localised near the unstable light ring at $ r = 3M $ (\textit{i.e.}, the so-called photon sphere). Later, Decanini, Folacci and Jensen showed that the complex frequencies of weakly damped quasinormal modes (QNMs) are Breit-Wigner resonances generated by the surface waves previously mentioned and, using the concept of Regge trajectories, they were able to construct semiclassically the spectrum of the QNM complex frequencies~\cite{Decanini:2002ha}. Recently, Folacci and Ould~El~Hadj have shown that CAM machinery can be used for precise numerical calculations of scattering cross sections of scalar fields, electromagnetic fields~\cite{Folacci:2019cmc} and gravitational waves~\cite{Folacci:2019vtt} on a Schwarzschild space-time. Moreover, using the third WKB approximation to obtain the asymptotic expressions for the lowest Regge poles~\cite{Decanini:2009mu} and associated residues, they have been able to provide an analytical approximation describing accurately both the ``glory'' and a large part of the ``orbiting'' oscillations in black hole scattering cross sections in the short-wavelength regime. An accurate approximation for the black hole absorption cross section has also been derived via an analysis of Regge poles~\cite{Decanini:2011xi,Decanini:2011xw}. In addition, an alternative description of gravitational radiation from black holes based on CAM theory was developed in Ref.~\cite{Folacci:2018sef}.

@@

There is a close relationship between Regge poles and QNM frequencies; both are sets of poles of the scattering matrix, with the latter lying in the complex-frequency plane. The excitation of black hole QNMs leads to distinctive signatures in \emph{time-dependent} scattering scenarios. For example, in the immediate aftermath of a black hole merger, the perturbed black hole returns to a quiescent state via a ``ringdown phase'' in which its gravitational-wave signal is well described as a sum of quasinormal modes \cite{Giesler:2019uxc}. The QNM spectrum of compact bodies have been well-studied by several authors  \cite{Detweiler:1985zz, Kokkotas:1986gd, Chandrasekhar449, Kokkotas:1992ka, Leins:1993zz, Andersson:1995ez, Andersson1996}, and are reviewed in Ref.~\cite{Kokkotas:1999bd}.

@@

%Finally we apply the CAM-approach developed for black hole scattering in \cite{Folacci:2019cmc} and \cite{Folacci:2019vtt}.

@@

Furthermore, since the wave equation is invariant under the transformation $\lambda \rightarrow -\lambda$, it follows that the distribution of poles in the left-half of the complex plane follows by reflection through the origin. Thus, for $\omega > 0$, the 4th and 2nd quadrants are devoid of poles, and the 3rd quadrant has the same structure as shown in Fig.~\ref{fig:rp1} and~\ref{fig:rp2}.
Conversely, for $\omega < 0$, and due to the symmetry relation~\eqref{symmetry_relation}, we may use $\lambda_n(-\omega) = \lambda_n(\omega)^*$ to establish that there are poles in the 2nd and 4th quadrants only.

@@

The Regge pole spectrum for a compact body is qualitatively similar to the Regge pole spectrum found in Mie scattering of electromagnetic waves by a transparent sphere of refractive index $\tilde{n}$. This has been studied since the 1960s; see for example Fig.~9.2 in Ref.~\cite{Nussenzveig:2006}. % In the latter case, the poles are given by the zeros of a combination of special functions.
Henceforth we shall adopt the terminology of Nussenzveig \cite{Nussenzveig:2006}, in which the three branches are labelled as:
\begin{enumerate}
 \item \emph{Broad resonances}: approximately uniformly-spaced poles above the real axis with approximately constant imaginary part; somewhat sensitive to internal structure ($r < R$);
 \item \emph{Surface waves}: highly-damped modes that are relatively insensitive to the internal structure and which depend chiefly on the surface geometry.
 \item \emph{Narrow resonances}: modes approaching the real axis corresponding to trapped modes which only appear in the UCO case ($R < 3M$).
\end{enumerate}
In Figs.~\ref{fig:rp1} and \ref{fig:rp2} the broad resonances, surface waves and narrow resonances are indicated by red squares, blue circles and purple triangles, respectively; and black diamonds indicate the black hole Regge poles. Whereas there exists an infinite number of poles of the surface-wave branch, in principle, there are no narrow-resonance poles at all in the $R/M = 6$ case, and just $\le 4$ narrow-resonance poles in the $R/M = 2.26$ case. The narrow-resonance branch is observed to end close to the start of the black hole branch.

@@

@@

To construct the scattering amplitude \eqref{Scalar_Scattering_amp}, the background integrals~\eqref{CAM_Scalar_Scattering_amp_decomp_Background_a} and~\eqref{CAM_Scalar_Scattering_amp_decomp_Background_b}, and the Regge pole amplitude~\eqref{CAM_Scalar_Scattering_amp_decomp_RP}, we use, \textit{mutatis mutandis}, the computational methods of Refs~\cite{Folacci:2019cmc,Folacci:2019vtt}. In these works the authors calculated the CAM representation of scattering amplitudes for scalar, electromagnetic and gravitational waves by Schwarzschild black hole (see also Ref~\cite{Dolan:2017rtj}).

Due to the long rang nature of the field propagating on the Schwarzschild black hole (outside the compact body), the scattering amplitude~\eqref{Scalar_Scattering_amp} and the background integral~\eqref{CAM_Scalar_Scattering_amp_decomp_Background_a} both suffer a lack of convergence. To overcome this problem, \textit{i.e.}, to accelerate the convergence of this sum and integral, we have used the method described in the Appendix of Ref~\cite{Folacci:2019cmc}. We have performed all the numerical calculations by using {\it Mathematica}. % \cite{Mathematica}.

@@

The CAM toolkit was applied here to gain a complementary method for calculating scattering amplitudes. In summary, we have shown that, for the intermediate and high angles, these scattering amplitudes can be reconstructed in terms of Regge poles with great precision in the intermediate and high frequency regimes. For compact objects ($R/M\sim6$), we observed that it is necessary to take into account the background integral contributions to describe the glory, and this is true regardless of the frequency. Conversely, this is not the case for UCOs ($R/M < 3$), where the sum over the Regge poles is sufficient to accurately describe the glory. Moreover, the sum over the Regge poles permitted us to overcome the difficulties linked to the lack of convergence from which the scattering amplitude suffers (\textit{i.e.} partial wave expansions) due to the long-range of the field propagating on the Schwarzschild spacetime outside the compact body. However, the existence of the broad resonance branch %(see also narrow resonance branch for the UCOs)
close to the real axis with an approximately uniform spacing, $\Delta\lambda\sim 2$, requires the calculation of a large number of poles. In the black hole case, by contrast, there is only a single branch of surface-wave-type modes (see Fig.~\ref{RP_R6_approx_2Mw_3_6_s_1}), and in the high-frequency limit, a few poles ($<5$) capture both the orbiting and glory phenomena \cite{Folacci:2019cmc, Folacci:2019vtt}.

@@

As discussed in Sec.~\ref{subsec:results1}, a good analytic extension of the S-matrix into the left-half complex $\lambda$-plane leads to the symmetry \eqref{Matrix_S_CAM_symm}. Then, it follows that the Regge poles with $\text{Re}[\lambda] < 0 $ may be found for (\textit{c.f.} \eqref{Evaluation_W_bis_1} and \eqref{Relation_1st_4th})
%
\begin{equation}
    \begin{dcases}
        \omega>0 \\
        \text{Im}[\lambda]<0
     \end{dcases}
\quad\text{or}\quad
    \begin{dcases}
      \omega<0 \\
        \text{Im}[\lambda]>0
    \end{dcases}.
\end{equation}\
